### 2022-01-25 plos_latex_template.tex
Tiny edits by @andreasancheztapia, @rociojoo + author affiliations

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*\Yinyang}, 
Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4},
Heather Turner\textsuperscript{5}, %\ddag},
Dorothea Hug Peter\textsuperscript{6}, %\ddag},
Matt Bannert\textsuperscript{7},
Batool Almazrouq\textsuperscript{8},
Liz Hare\textsuperscript{9},
Laura Ación\textsuperscript{10},
Juan Pablo Narváez-Gómez\textsuperscript{11},
Marcela Alfaro Córdoba\textsuperscript{12},
Federico Marini\textsuperscript{13},
Rita Giordano\textsuperscript{14},
Natalia Morandeira\textsuperscript{15},
Adithi Upadhya\textsuperscript{16},
Joselyn Chávez\textsuperscript{17},
Anicet Ebou\textsuperscript{18},
Silvia Canelón\textsuperscript{19},
Janani Ravi\textsuperscript{20}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{4} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{5} Department of Statistics, University of Warwick, Coventry, UK
\\
\textbf{6} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{7} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{8} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{9} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{10} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{11} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{12} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{13} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{14} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{15} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{16} ILK Labs/Bengaluru, India - 560046
\\
\textbf{17} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{18} Bioinformatic team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
\\
\textbf{19} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{20} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
\\
\bigskip

@@

Conferences are spaces to meet and network within an academic or technical field, to learn about new advances and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces 10 simple rules to host an inclusive conference. 
The authors recently organized the 2021 edition of the useR! statistical computing conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been mainly written for potential organizers or selection committees of conferences, and contain multiple practical tips so that they can help diverse types of scientific and technological conferences to become as accessible and inclusive as they can and should be.

@@

% Conferences are great, but only for some
Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
A good conference experience can make a difference in the professional development of the participants and create long-lasting collaborations and opportunities. 
However, opportunities for participating in conferences are not equal for all. 
Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from rich countries, prestigious institutions, native English speakers with no physical disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
The authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
We had the opportunity to organize a virtual and global statistical computing conference, useR! 2021 \cite{sanchez-tapia_user_2021-2}, for users and developers of the R programming language \cite{r_core_team_2021}.
Coming from different backgrounds, career stages and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.

% Roadmap of the paper
The rules result from our lessons learned before, during, and after the organization of the useR! conference. 
They are organized in three sections: the pillars, design rules, and the continuity rule (\textbf{Fig. \ref{fig:diagram}}).
\textbf{The pillars} comprise three fundamental rules with key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
Six \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and other spaces). 
\textbf{Rule 5} provides advice for organizing inclusive online and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities. 
In \textbf{Rule 7}, we provide suggestions to encourage and ensure full participation of non-native English speakers. 
\textbf{Rule 8} provides tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
a \textbf{continuity Rule} to make the conference part of a long-term commitment to inclusion.

@@

% A culture of support (active bystandership) and code of conduct for a safe place 
Creating a safe and welcoming environment also requires promoting a culture of support, including the idea that anyone can intervene respectfully in defense of other attendees in case anything unpleasant, incorrect, or offensive happens.
This is called `active bystandership'; consider providing training or pre-conference discussions to promote it.
However, the most important step that you should take to ensure the safety of the community is
adopting a code of conduct (coc) and preparing a diverse response team to enforce it \cite{favaroYourScienceConference2016}.
The coc is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behaviors, and the way to report violations \cite{auroraHowRespondCode2019}.
The coc response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community. The team should be prepared early during conference planning because coc violations can happen before the conference too (e.g. social media, organization meetings).  
We strongly recommend reading `How to Respond to code of conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong coc team.

@@

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
The regional and local communities, groups, or associations in your field are good sources to tap into. 
When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
People from underrepresented groups sometimes refrain to nominate themselves while people in dominant groups tend to do it more easily. 
Instead, nominate folks directly and offer them leading positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget (see \textbf{Rule 9}).
Furthermore, tasks such as receiving and responding to code of conduct reports can be emotionally intense work and should be additionally rewarded.
And most importantly, take care of your team and their well-being! Check on them regularly and make sure that everyone is comfortable. This represents a great amount of work and you will need to support each other in the long run.

@@

Biases in the final list of abstracts begin with self-selection: some potential conference participants may not feel confident enough to submit their work, especially to large and prestigious conferences.
To mitigate this self-selection bias, 
some communities have pre-submission mechanisms to aid their members prepare and receive feedback for their abstract (e.g. the Latin-R and R-Ladies communities have volunteers to help their members improve their conference abstracts).
If this does not exist in your field yet, consider asking members of your community to provide a similar kind of mentorship.

@@

In many academic or technological fields, typically competitive activities or areas—such as scientific publications or software development—are more valued than those focused on sharing and cooperation, like community building, teaching or mentoring. 
The latter activities are equally, if not more, relevant and challenging, and are usually performed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020, light_gender_2022}.
When planning the program for the conference, consider giving visibility to the whole range of activities and practitioners that contribute to the field by proposing new thematic sessions, broadening the scope of talks, keynotes, and tutorials.

@@

If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and language interpretation, to name a few. 
Invisible disabilities should be accounted for proactively (e.g. dyslexia, anxiety, ADHD, autism), for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.

Regardless of the conference format, all platforms (website, registration, abstract submission, chat system, conference administration tools, live streaming) should be screen-reader friendly and keyboard accessible. 
All images used in digital conference spaces should have alternative text and use color-blind safe palettes.
If there is video streaming, it should have good-quality (not automated) captions and a transcript; live presentations should have good-quality live captions too.
If the conference has a regional scope, sign-language interpreters can be a better option than captions.

@@

The communication strategy of your conference should aim to reach broader audiences, express the commitment with diversity and inclusion, and the welcoming spirit of the conference.
Actively reach out and promote the conference to communities that have been systematically excluded. 
Promote the abstract submission call beyond the usual communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
The members of your team should be able to decide which languages to emphasize and which social media platforms to utilize for the promotion of the conference (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists).%#promotion2

@@

\begin{figure}[!h]
\centering
\pdftooltip{
\includegraphics[width=\textwidth]{figs/marmots.pdf}}{
}
\caption{Margot the marmot was the useR! 2021 mascot. To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing the colors of diverse communities of practice that are part of the large R community and specific groups of people we wanted to support explicitly: (from left to right) scarves with MiR (Minorities in R https://mircommunity.com/), R-Ladies https://rladies.org/, AfricaR https://africa-r.org/, rainbow LGBTQIA+ and transgender colors. Margot was created by Francisco Etchart and is available under a CC-BY-NC-SA license} 
\label{fig:marmots}
\end{figure}
%alt-text for the figure, to be included in the final PDF: Five marmots with colorful scarves. From left to right: a marmot with a dark and light green scarf from the MiR community is watching her tablet, a marmot with the purple and gray scarf for R-Ladies is stepping on a rock and watches the horizon, a marmot with an orange and purple AfricaR scarf is sitting on a couch watching a film on her laptop with popcorn, a yogi marmot with a rainbow LGBTQIA+ scarf does the warrior pose on her blue mat, and finally, a marmot with a scarf with the pink, light blue and white color of the transgender pride flag waves her hand. 
\subsection*{Rule 9: Allocate adequate financial resources to support inclusion goals}
\label{rule_financial}

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
When determining the registration rates, the socioeconomic context of participants, their country of origin, and their career status should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
Resources permitting, enable a `pay what you can' option. You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}. 
Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
Conferences usually ask for cover letters or applications to assign these scholarships, which can be time-consuming and emotionally demanding. 
Simplifying the process of requesting financial support, especially for 
people who lack the time and resources (e.g. because of family responsibilities) could greatly increase their chances of applying for the support. 
Additional aid for attendees could also be offered—e.g. child care support or internet connection services for the virtual component. 
Bear in mind that transferring money internationally could be a cumbersome administrative process and significantly augment the burden of already marginalized groups. Whenever possible, facilitate alternate ways for money transfer (e.g. book flights, hotel reservations, waive conference fees).

@@

Be aware of all the factors that may hinder your efforts for inclusion: systemic discrimination, time or money constraints, geopolitical or public health contexts, or personal issues affecting your organizing team. 
Try to account for these factors as best as you can to set realistic yet bold vision and goals of diversity and inclusion for the conference.
At the end of the day, even if you do not achieve them at 100\%, striving for them will make you stretch what feels possible and help you make changes you would have otherwise not even attempted to make. 
And most importantly, you would have contributed to a great conference experience for the attendees. 
They will notice the welcoming and inclusive spirit of the conference reflected in the pillar rules, and translated into design decisions—even if there are gaps in them due to the constraining factors mentioned above. 
Inclusive conferences can help to enhance career paths and create a sense of belonging in the community. 
That is more than enough motivation to make the effort worth it.

@@

\begin{itemize}
    \item RJ: Project administration, Supervision, Writing original draft, Conceptualization, Writing--major reviews and edits, Resources, Visualization
    \item AST: Project administration, Supervision, Conceptualization,  Writing--major reviews and edits, Resources, Visualization, Sharing rules
    \item SM: Conceptualization, Writing--major reviews and edits, Resources, Visualization, Sharing rules
    \item YBS: Conceptualization, Writing--major reviews and edits, Resources
    \item HT: Conceptualization, Writing--major reviews and edits, Resources
    \item DHP: Conceptualization, Writing--major reviews and edits, Sharing rules
    \item MB: Conceptualization
    \item BA: Writing--major reviews and edits, Resources, Sharing rules
    \item LH: Writing--major reviews and edits, Resources, Sharing rules
    \item LA: Writing--major reviews and edits, Resources
    \item JPN: Writing--major reviews and edits, Visualization
    \item MAC: Writing--major reviews and edits
    \item FM: Writing--major reviews and edits
    \item RG: Writing--major reviews and edits
    \item NM: Sharing rules
    \item AU: Sharing rules
    \item JC: Sharing rules
    \item AE: Sharing rules
    \item SC: Sharing rules
    \item JR: Writing--major reviews and edits
\end{itemize}

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*\Yinyang}, 
Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4},
Heather Turner\textsuperscript{5}, %\ddag},
Dorothea Hug Peter\textsuperscript{6}, %\ddag},
Matt Bannert\textsuperscript{7},
Batool Almazrouq\textsuperscript{8,9},
Liz Hare\textsuperscript{10},
Laura Ación\textsuperscript{11},
Juan Pablo Narváez-Gómez\textsuperscript{12},
Marcela Alfaro Córdoba\textsuperscript{13},
Federico Marini\textsuperscript{14},
Rita Giordano\textsuperscript{15},
Natalia Morandeira\textsuperscript{16},
Adithi Upadhya\textsuperscript{17},
Joselyn Chávez\textsuperscript{18},
Anicet Ebou\textsuperscript{19},
Silvia Canelón\textsuperscript{20},
Janani Ravi\textsuperscript{21}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{4} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{5} Department of Statistics, University of Warwick, Coventry, UK
\\
\textbf{6} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{7} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{8} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK
\\
\textbf{9} King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
\\
\textbf{10} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{11} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{12} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{13} Department of Statistics, University of California, Santa Cruz, CA, USA
\\
\textbf{14} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{15} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{16} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{17} ILK Labs/Bengaluru, India - 560046
\\
\textbf{17} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{19} Bioinformatic team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
\\
\textbf{20} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{21} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
\\
\bigskip

@@

Conferences are spaces to meet and network within an academic or technical field, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces 10 simple rules to host an inclusive conference. 
The authors recently organized the 2021 edition of the useR! statistical computing conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been mainly written for potential organizers or selection committees of conferences, and contain multiple practical tips so that they can help diverse types of events to become as accessible and inclusive as they can and should be.

@@

% Conferences are great, but only for some
Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
A good conference experience can make a difference in the professional development of the participants and create long-lasting collaborations and opportunities. 
However, opportunities for participating in conferences are not equal for all. 
Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from rich countries, prestigious institutions, native English speakers with no physical disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
The authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
We had the opportunity to organize a virtual and global statistical computing conference, useR! 2021 \cite{sanchez-tapia_user_2021-2}, for users and developers of the R programming language \cite{r_core_team_2021}.
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.

% Roadmap of the paper
The rules result from our lessons learned before, during, and after the organization of the useR! conference. 
They are organized in three sections: three pillars, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
\textbf{The pillars} comprise fundamental rules with key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and other spaces). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to encourage and ensure full participation of non-native English speakers. 
\textbf{Rule 8} provides tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
a \textbf{continuity Rule} emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion.

@@

% A culture of support (active bystandership) and code of conduct for a safe place 
Creating a safe and welcoming environment also requires promoting a culture of support, including the idea that anyone can intervene respectfully in defense of other attendees in case anything unpleasant, incorrect, or offensive happens.
This is called `active bystandership'; consider providing training or pre-conference discussions to promote it.
However, the most important step that you should take to ensure the safety of the community is
adopting a code of conduct (CoC) and preparing a diverse response team to enforce it \cite{favaroYourScienceConference2016}.
The CoC is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behaviors, and the way to report violations \cite{auroraHowRespondCode2019}.
The CoC response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community. The team should be prepared early during conference planning because CoC violations can happen before the conference too (e.g. social media, organization meetings).  
We strongly recommend reading `How to Respond to code of conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC response team.

@@

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
The regional and local communities, groups, or associations in your field are good sources to tap into. 
When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
People from underrepresented groups sometimes refrain to nominate themselves while people in dominant groups tend to do it more easily. 
Instead, nominate folks directly and offer them leading positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider paying them as an item in your budget (see \textbf{Rule 9}).
Furthermore, tasks such as receiving and responding to code of conduct reports can be emotionally intense work and should be additionally rewarded.
And most importantly, take care of your team and their well-being! Check on them regularly and make sure that everyone is comfortable. This represents a great amount of work and you will need to support each other in the long run.

@@

Biases in the final list of abstracts begin with self-selection: some potential conference participants may not feel confident enough to submit their work, especially to large and prestigious conferences.
To mitigate this, 
some communities have pre-submission mechanisms to aid their members prepare and receive feedback for their abstract (e.g. the Latin-R and R-Ladies communities have volunteers to help their members improve their conference abstracts).
If this does not exist in your field yet, consider asking members of your community to provide a similar kind of mentorship.

@@

In many academic or technological fields, typically competitive activities or areas—such as scientific publications or software development—are more valued than those focused on sharing and cooperation, like community building, teaching, or mentoring. 
The latter activities are equally, if not more, relevant and challenging, and are usually performed by women, racialized people, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020, light_gender_2022}.
When planning the program for the conference, consider giving visibility to the whole range of activities and practitioners that contribute to the field by proposing new thematic sessions, broadening the scope of talks, keynotes, and tutorials.

@@

If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and language interpretation, to name a few. 
Invisible disabilities  (e.g. dyslexia, anxiety, ADHD, autism) should be accounted for proactively, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.

Regardless of the conference format, all platforms (website, registration, abstract submission, chat system, conference administration tools, live streaming) should be screen-reader friendly and keyboard accessible. 
All images used in digital conference spaces should have alternative text and use colorblind-safe palettes.
If there is video streaming, it should have good-quality (not automated) captions and a transcript; live presentations should have good-quality live captions too.
If the conference has a regional scope, sign language interpreters can be a better option than captions.

@@

The communication strategy of your conference should aim to reach broader audiences and express the commitment with diversity and inclusion and the welcoming spirit of the conference.
Actively reach out and promote the conference to communities that have been systematically excluded. 
Promote the abstract submission call beyond the usual communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
The members of your team should be able to decide which languages to emphasize and which social media platforms to utilize for the promotion of the conference (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists).%#promotion2

@@

\begin{figure}[!h]
\centering
\pdftooltip{
\includegraphics[width=\textwidth]{figs/marmots.pdf}}{
}
\caption{Margot the marmot was the useR! 2021 mascot. To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing the colors of diverse communities of practice that are part of the large R community and specific groups of people we wanted to support explicitly. From left to right: Margot wearing scarves with the colors of MiR (Minorities in R https://mircommunity.com/), R-Ladies https://rladies.org/, AfricaR https://africa-r.org/, and the LGBTQIA+ and transgender communities. Margot was created by Francisco Etchart and is available under a CC-BY-NC-SA license} 
\label{fig:marmots}
\end{figure}
%alt-text for the figure, to be included in the final PDF: Five marmots with colorful scarves. From left to right: a marmot with a dark and light green scarf from the MiR community is watching her tablet, a marmot with the purple and gray scarf for R-Ladies is stepping on a rock and watches the horizon, a marmot with an orange and purple AfricaR scarf is sitting on a couch watching a film on her laptop with popcorn, a yogi marmot with a rainbow LGBTQIA+ scarf does the warrior pose on her blue mat, and finally, a marmot with a scarf with the pink, light blue and white color of the transgender pride flag waves her hand.

\subsection*{Rule 9: Allocate adequate financial resources to support inclusion goals}
\label{rule_financial}

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
When determining the registration rates, the socioeconomic context of participants, their country of origin, and their career status should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
Resources permitting, enable a `pay what you can' option. You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}. 
Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
To assign these scholarships, conferences usually ask for cover letters or applications, which can be time-consuming and emotionally demanding. 
Simplifying the process of requesting financial support, especially for 
people who lack the time and resources (e.g. because of family responsibilities) could greatly increase their chances of applying for the support. 
Additional aid for attendees could also be offered—e.g. paying for child care or internet connection services for the virtual component. 
Bear in mind that transferring money internationally could be a cumbersome administrative process and significantly augment the burden of already marginalized groups. Whenever possible, facilitate alternate ways for money transfer (e.g. book flights, hotel reservations, waive conference fees, give gift cards). Depending on the country, services like OpenCollective (https://opencollective.com/) can be very useful to do these transfers.

@@

Be aware of all the factors that may hinder your efforts for inclusion: systemic discrimination, time or money constraints, geopolitical or public health contexts, or personal issues affecting your organizing team. 
Try to account for these factors as best as you can to set a realistic yet bold vision and goals of diversity and inclusion for the conference.
At the end of the day, even if you do not achieve them at 100\%, striving for them will make you stretch what feels possible and help you make changes you would have otherwise not even attempted to make. 
And most importantly, you would have contributed to a great conference experience for the attendees. 
They will notice the welcoming and inclusive spirit of the conference reflected in the pillar rules and the effort put in every aspect of the conference design. 
Inclusive conferences can have lasting impacts in career paths and create a sense of belonging in the community. 
That is more than enough motivation to make the effort worth it.

@@

\begin{itemize}
    \item RJ: Project administration, Supervision, Writing original draft, Conceptualization, Writing--major reviews and edits, Resources, Visualization
    \item AST: Project administration, Supervision, Conceptualization,  Writing--major reviews and edits, Resources, Visualization, Sharing rules
    \item SM: Conceptualization, Writing--major reviews and edits, Resources, Visualization, Sharing rules
    \item YBS: Conceptualization, Writing--major reviews and edits, Resources
    \item HT: Conceptualization, Writing--major reviews and edits, Resources
    \item DHP: Conceptualization, Writing--major reviews and edits, Sharing rules
    \item MB: Conceptualization
    \item BA: Writing--major reviews and edits, Resources, Sharing rules
    \item LH: Writing--major reviews and edits, Resources, Sharing rules
    \item LA: Writing--major reviews and edits, Resources
    \item JPN: Writing--major reviews and edits, Visualization
    \item MAC: Writing--major reviews and edits
    \item FM: Writing--major reviews and edits
    \item RG: Writing--major reviews and edits
    \item NM: Sharing rules
    \item AU: Sharing rules
    \item JC: Sharing rules
    \item AE: Sharing rules
    \item SC: Sharing rules
    \item JR: Conceptualization, Writing--major reviews and edits
\end{itemize}
